---
title: "GARCH and PIT"
output: html_document
date: "2026-03-10"
---


```{r}
#load the data
setwd("C:/Users/Laopinou/Desktop/thesis")
library(readr)
library(rugarch)
library(readr)
library(dplyr)
library(VineCopula)
library(network)
data <- read_csv("weekly_prices_balanced.csv")
```


Since weekly_prices_balanced only contains weekly adjusted closing prices, we need to convert them into weekly returns first. 
```{r}
# Compute simple returns
price_matrix <- data[ , -1]
return_matrix <- price_matrix[-1, ] / price_matrix[-nrow(price_matrix), ] - 1
returns <- data.frame(
  Date = data$Date[-1],
  return_matrix
)

summary(returns[,-1])
```

Now with the weekly returns, we can fit the ARMA-GARCH model. we choose to fit AR(1)-GARCH(1,1) for all of the 30 asset returns.
```{r}

#ARMA-GARCH
returns <- returns %>% arrange(Date)
asset_cols <- setdiff(names(returns), "Date")

#Define the AR(1)-GARCH(1,1)-t specification
spec <- ugarchspec(
  variance.model = list(model = "sGARCH", garchOrder = c(1, 1)),
  mean.model     = list(armaOrder = c(1, 0), include.mean = TRUE),
  distribution.model = "std"   #Student-t
)

#fit the model
fits <- list()
std_resid <- data.frame(Date = returns$Date)
sigma_hat <- data.frame(Date = returns$Date)
mu_hat    <- data.frame(Date = returns$Date)

for (sym in asset_cols) {
  x <- returns[[sym]]
  
  # Fit (remove NAs just in case)
  fit <- ugarchfit(spec = spec, data = x, solver = "hybrid")
  fits[[sym]] <- fit
  
  # Extract series (same length as input)
  std_resid[[sym]] <- as.numeric(residuals(fit, standardize = TRUE))
  sigma_hat[[sym]] <- as.numeric(sigma(fit))
  mu_hat[[sym]]    <- as.numeric(fitted(fit))
}

```




```{r}
#check asset summary("SPY" for example)
show(fits[["SPY"]])
```

We need to test if all the residuals are approximately which noise. 
```{r}
asset_cols <- setdiff(names(std_resid), "Date")
Z <- std_resid[asset_cols]

#serial correlation check 
lb_resid <- sapply(Z, function(x){
  Box.test(x, lag = 20, type = "Ljung-Box")$p.value
})

lb_resid
sum(lb_resid < 0.05)

#Ljung-Box test for squared residuals to check ARCH effects
lb_sq <- sapply(Z, function(x){
  Box.test(x^2, lag = 20, type = "Ljung-Box")$p.value
})

lb_sq
sum(lb_sq < 0.05)
```


As shown in the results, at 0.05 level,there is no remaining serial dependence nor ARCH effects.


Now we also want to check if innovation is correctly specified as our predefined t-distribution
```{r}
# estimated degrees of freedom for each asset
df_vec <- sapply(asset_cols, function(sym) {
  as.numeric(coef(fits[[sym]])["shape"])
})

names(df_vec) <- asset_cols
df_vec
Z <- std_resid[asset_cols]
par(mfrow = c(2, 2), mar = c(4, 4, 2, 1))

for (sym in asset_cols) {
  z  <- Z[[sym]]
  z  <- z[is.finite(z)]
  nu <- as.numeric(df_vec[sym])
  
  if (length(z) < 10) {
    message("Skipping ", sym, ": too few finite residuals")
    next
  }
  
  if (!is.finite(nu) || is.na(nu) || nu <= 0) {
    message("Skipping ", sym, ": invalid df")
    next
  }
  
  p <- ppoints(length(z))
  q_theo <- qt(p, df = nu)
  q_emp  <- sort(z)
  
  plot(q_theo, q_emp,
       main = paste("QQ:", sym),
       xlab = "Theoretical t quantiles",
       ylab = "Empirical residual quantiles")
  abline(0, 1, col = "red", lwd = 2)
}

```


The next step is Probability Integral Transform(PIT) with parametric t-distribution
```{r}

##parametric PIT
asset_cols <- setdiff(names(std_resid), "Date")

df_vec <- sapply(asset_cols, function(sym) {
  as.numeric(coef(fits[[sym]])["shape"])
})

names(df_vec) <- asset_cols
df_vec

U_t <- data.frame(Date = std_resid$Date)

for (sym in asset_cols) {
  z  <- std_resid[[sym]]
  nu <- df_vec[sym]
  
  u <- rep(NA_real_, length(z))
  ok <- is.finite(z) & is.finite(nu) & !is.na(nu) & nu > 0
  
  u[ok] <- pt(z[ok], df = nu)
  
  U_t[[sym]] <- u
}

U_t_mat <- as.matrix(U_t[asset_cols])

```



Fitting a multivariate t-copula as a baseline model 
```{r}
library(copula)
d <- ncol(U_t_mat)
t_cop <- tCopula(dim = d, dispstr = "un")
fit_t <- fitCopula(t_cop, data = U_t_mat, method = "mpl")
saveRDS(fit_t, "fit_t_copula_ut.rds")

```


```{r}
readRDS("fit_t_copula_ut.rds")
```





R-vine copula with truncated Trees 
```{r}
#define the elliptical and Archimedian families
family_set <- c(
  1,   # Gaussian
  2,   # Student t
  3,   # Clayton
  4,   # Gumbel
  5,   # Frank
  6,   # Joe
  13,  # survival Clayton
  14,  # survival Gumbel
  16   # survival Joe
)


```



```{r}
#R vine with paramtric PIT
#independence=false

Utd<-U_t_mat
max_K <- ncol(Utd) - 1
bic_vals <- rep(NA, max_K)
fit_rvine_td <- vector("list", max_K)



time_loop <- system.time({
for (K in 1:max_K) {
  fit_rvine_td[[K]] <- RVineStructureSelect(
    data = Utd,
    familyset = family_set,
    selectioncrit = "AIC",   # pair-copula family choice
    indeptest = FALSE,
    trunclevel = K
  )
  
  bic_vals[K] <- RVineBIC(Utd, fit_rvine_td[[K]])
  cat("Finished K =", K, "\n")
}
})

time_loop
time_loop["elapsed"]

bic_vals
bic_vals<-as.numeric(bic_vals)
delta_bic <- diff(bic_vals)
delta_bic
```

We can see that the BIC no longer decrease after tree 7, thus our truncation level is set at Tree 7. We save the results of the  fianl truncated model (at K=7) in an rds file to save time.

```{r}
rvine_td=fit_rvine_td[[7]]
rvine_td

```

```{r}
saveRDS(rvine_td, "rvine_td.rds")
```


We try to draw the structure of the first tree
```{r}
plot(
  rvine_td,
  tree = 1,
  vertex.cex = 0.8,   # VERY IMPORTANT
  label.cex = 1,
)
```

```{r}

#another way to draw the graph 
library(igraph)

# Extract vine matrix
M <- rvine_td$Matrix
d <- nrow(M)

# Tree 1 variable order
order_tree1 <- rev(M[d, ])

# Consecutive edges in Tree 1
edge_list <- cbind(order_tree1[-length(order_tree1)],
                   order_tree1[-1])

# Build graph
g <- graph_from_edgelist(edge_list, directed = FALSE)

# Plot
plot(
  g,
  layout = layout_with_fr,
  vertex.size = 28,
  vertex.color = "lightblue",
  vertex.label.cex = 1.1,
  edge.width = 2
)
```



Now we so R-vine simulation with truncation level K=7
```{r}
#prepare data
returns <- returns %>% arrange(Date)
asset_cols <- setdiff(names(returns), "Date")

# equal-weight portfolio
w <- rep(1 / length(asset_cols), length(asset_cols))
names(w) <- asset_cols

# AR(1)-GARCH(1,1)-t specification
spec <- ugarchspec(
  variance.model = list(model = "sGARCH", garchOrder = c(1, 1)),
  mean.model     = list(armaOrder = c(1, 0), include.mean = TRUE),
  distribution.model = "std"
)

```












For theR vine simulation, from the previous truncated R vine, we got the truncation level=6
```{r}
set.seed(123)
```

```{r}
# 1. Rolling window settings
#-----------------------------
W <- 500                  # rolling window length
B <- 10000              # number of Monte Carlo simulations
K_trunc <- 6              # truncate R-vine at Tree 6

Tn <- nrow(returns)
n_oos <- Tn - W          # number of out-of-sample forecasts

# storage
backtest_results <- data.frame(
  Date = returns$Date[(W + 1):Tn],
  VaR_95 = NA_real_,
  VaR_99 = NA_real_,
  Rp_real = NA_real_,
  Violation_95 = NA_integer_,
  Violation_99 = NA_integer_
)

# store fitted vine objects if you want
rvine_list <- vector("list", n_oos)
```

```{r}
# 2. Rolling backtest loop
#-----------------------------
for (tt in W:(Tn - 1)) {
  
  cat("Rolling step:", tt, "of", Tn - 1, "\n")
  
  #-------------------------
  # A. Define train / test
  #-------------------------
  train_idx <- (tt - W + 1):tt
  test_idx  <- tt + 1
  
  train_data <- returns[train_idx, ]
  test_data  <- returns[test_idx, ]
  
  #-------------------------
  # B. Fit AR-GARCH marginals
  #-------------------------
  fits <- list()
  std_resid <- matrix(NA_real_, nrow = W, ncol = length(asset_cols))
  colnames(std_resid) <- asset_cols
  
  mu_next <- numeric(length(asset_cols))
  sigma_next <- numeric(length(asset_cols))
  df_vec <- numeric(length(asset_cols))
  
  names(mu_next) <- asset_cols
  names(sigma_next) <- asset_cols
  names(df_vec) <- asset_cols
  
  fit_ok <- TRUE
  
  for (sym in asset_cols) {
    x <- train_data[[sym]]
    
    fit <- tryCatch(
      ugarchfit(spec = spec, data = x, solver = "hybrid"),
      error = function(e) NULL
    )
    
    if (is.null(fit)) {
      fit_ok <- FALSE
      cat("GARCH failed for", sym, "at step", tt, "\n")
      break
    }
    
    fits[[sym]] <- fit
    
    # in-sample standardized residuals
    std_resid[, sym] <- as.numeric(residuals(fit, standardize = TRUE))
    
    # one-step-ahead forecasts
    fcst <- tryCatch(
      ugarchforecast(fit, n.ahead = 1),
      error = function(e) NULL
    )
    
    if (is.null(fcst)) {
      fit_ok <- FALSE
      cat("Forecast failed for", sym, "at step", tt, "\n")
      break
    }
    
    mu_next[sym] <- as.numeric(fitted(fcst))
    sigma_next[sym] <- as.numeric(sigma(fcst))
    df_vec[sym] <- as.numeric(coef(fit)["shape"])
  }
  
  if (!fit_ok) next
  
  #-------------------------
  # C. Parametric PIT
  #-------------------------
  U_t_mat <- matrix(NA_real_, nrow = W, ncol = length(asset_cols))
  colnames(U_t_mat) <- asset_cols
  
  for (sym in asset_cols) {
    z <- std_resid[, sym]
    nu <- df_vec[sym]
    
    ok <- is.finite(z) & is.finite(nu) & !is.na(nu) & nu > 0
    u <- rep(NA_real_, length(z))
    u[ok] <- pt(z[ok], df = nu)
    
    # avoid exact 0 or 1
    eps <- 1e-6
    u[u <= eps] <- eps
    u[u >= 1 - eps] <- 1 - eps
    
    U_t_mat[, sym] <- u
  }
  
  # remove rows with NA just in case
  complete_rows <- complete.cases(U_t_mat)
  Utd <- U_t_mat[complete_rows, , drop = FALSE]
  
  if (nrow(Utd) < 50) {
    cat("Too few complete PIT rows at step", tt, "\n")
    next
  }
  
  #-------------------------
  # D. Fit truncated R-vine
  #-------------------------
  family_set <- c(
  1,   # Gaussian
  2,   # Student t
  3,   # Clayton
  4,   # Gumbel
  5,   # Frank
  6,   # Joe
  13,  # survival Clayton
  14,  # survival Gumbel
  16   # survival Joe
)

  
  fit_rvine <- tryCatch(
    RVineStructureSelect(
      data = Utd,
      familyset = family_set,      
      selectioncrit = "AIC",
      indeptest = FALSE,
      trunclevel = K_trunc
    ),
    error = function(e) NULL
  )
  
  if (is.null(fit_rvine)) {
    cat("R-vine fit failed at step", tt, "\n")
    next
  }
  
  rvine_list[[tt - W + 1]] <- fit_rvine
  
  #-------------------------
  # E. Simulate from fitted R-vine
  #-------------------------
  U_sim <- tryCatch(
    RVineSim(N = B, RVM = fit_rvine),
    error = function(e) NULL
  )
  
  if (is.null(U_sim)) {
    cat("R-vine simulation failed at step", tt, "\n")
    next
  }
  
  #-------------------------
  # F. Transform uniforms to innovations
  #    using marginal t inverse CDF
  #-------------------------
  d <- length(asset_cols)
  Z_sim <- matrix(NA_real_, nrow = B, ncol = d)
  colnames(Z_sim) <- asset_cols
  
  for (i in seq_along(asset_cols)) {
    sym <- asset_cols[i]
    Z_sim[, i] <- qt(U_sim[, i], df = df_vec[sym])
  }
  
  #-------------------------
  # G. Convert innovations to returns
  #-------------------------
  R_sim <- matrix(NA_real_, nrow = B, ncol = d)
  colnames(R_sim) <- asset_cols
  
  for (i in seq_along(asset_cols)) {
    sym <- asset_cols[i]
    R_sim[, i] <- mu_next[sym] + sigma_next[sym] * Z_sim[, i]
  }
  
  #-------------------------
  # H. Simulated portfolio returns
  #-------------------------
  port_ret_sim <- as.vector(R_sim %*% w)
  
  # VaR as positive loss numbers
  VaR_95_t <- -as.numeric(quantile(port_ret_sim, probs = 0.05, na.rm = TRUE, type = 8))
  VaR_99_t <- -as.numeric(quantile(port_ret_sim, probs = 0.01, na.rm = TRUE, type = 8))
  
  #-------------------------
  # I. Realized next-period portfolio return
  #-------------------------
  Rp_real_t <- sum(as.numeric(test_data[1, asset_cols]) * w)
  
  #-------------------------
  # J. Record violations
  #-------------------------
  row_out <- tt - W + 1
  
  backtest_results$VaR_95[row_out] <- VaR_95_t
  backtest_results$VaR_99[row_out] <- VaR_99_t
  backtest_results$Rp_real[row_out] <- Rp_real_t
  backtest_results$Violation_95[row_out] <- as.integer(Rp_real_t < -VaR_95_t)
  backtest_results$Violation_99[row_out] <- as.integer(Rp_real_t < -VaR_99_t)
}

```
```{r}
saveRDS(backtest_results,"backtest_results_td.rds")
```

```{r}
backtest_results_td <- readRDS("backtest_results_td.rds")

```

```{r}
# remove failed rows if any
bt_td <- backtest_results_td %>%
  filter(!is.na(VaR_95), !is.na(Rp_real))


# empirical violation rates
viol_rate_95 <- mean(bt_td$Violation_95)
viol_rate_99 <- mean(bt_td$Violation_99)

viol_rate_95
viol_rate_99

# expected rates
cat("95% VaR expected violation rate: 0.05\n")
cat("95% VaR observed violation rate:", viol_rate_95, "\n\n")

cat("99% VaR expected violation rate: 0.01\n")
cat("99% VaR observed violation rate:", viol_rate_99, "\n")
```
```{r}
kupiec_test <- function(violations, alpha) {
  x <- sum(violations)
  n <- length(violations)
  p_hat <- x / n
  
  # handle edge cases
  if (p_hat == 0 || p_hat == 1) {
    return(list(
      LRuc = NA,
      p_value = NA,
      violations = x,
      n = n,
      observed_rate = p_hat
    ))
  }
  
  LRuc <- -2 * (
    (n - x) * log(1 - alpha) + x * log(alpha) -
      ((n - x) * log(1 - p_hat) + x * log(p_hat))
  )
  
  p_value <- 1 - pchisq(LRuc, df = 1)
  
  list(
    LRuc = LRuc,
    p_value = p_value,
    violations = x,
    n = n,
    observed_rate = p_hat
  )
}

kupiec_95 <- kupiec_test(bt_td$Violation_95, alpha = 0.05)
kupiec_99 <- kupiec_test(bt_td$Violation_99, alpha = 0.01)

kupiec_95
kupiec_99
```

```{r}
plot(
  bt_td$Date, bt_td$Rp_real, type = "l",
  main = "Realized Portfolio Returns vs VaR Forecasts:Truncated R-vine",
  xlab = "Date", ylab = "Realized Returns"
)
lines(bt_td$Date, -bt_td$VaR_95, lty = 1, lwd=1.5,col="blue")
lines(bt_td$Date, -bt_td$VaR_99, lty = 1, lwd=1.5,col="red")
legend(
  "bottomleft",
  legend = c("Realized portfolio return", "-VaR 95%", "-VaR 99%"),
  lty = c(1, 1, 1),
  col = c("black", "blue", "red"),
  bty = "n"
)
```



























C-vine Copula modeling with K=6
```{r}
# 1. Define asset class groups

groups <- list(
  Equity       = 1:10,
  FixedIncome  = 11:16,
  Commodities  = 17:21,
  Currencies   = 22:26,
  Alternatives = 27:30
)

# 2. Compute empirical Kendall's tau matrix

tau_mat <- cor(U_t_mat, method = "kendall", use = "pairwise.complete.obs")

```

```{r}
# Choose global hub v1
global_scores <- rowSums(abs(tau_mat)) - diag(abs(tau_mat))
global_hub_index <- which.max(global_scores)
global_hub_name  <- colnames(U_t_mat)[global_hub_index]

cat("Global hub v1:\n")
cat("Index:", global_hub_index, "\n")
cat("Name :", global_hub_name, "\n")
cat("Score:", global_scores[global_hub_index], "\n\n")

```
```{r}
# Class representatives excluding global hub

class_reps_excl_global <- data.frame(
  Class = character(),
  Index = integer(),
  Name  = character(),
  WithinClassScore = numeric(),
  GlobalScore = numeric(),
  stringsAsFactors = FALSE
)

for (gname in names(groups)) {
  idx <- groups[[gname]]
  
  # exclude global hub from representative selection
  idx2 <- setdiff(idx, global_hub_index)
  
  tau_sub <- tau_mat[idx2, idx2, drop = FALSE]
  scores_sub <- rowSums(abs(tau_sub)) - diag(abs(tau_sub))
  
  rep_local_pos <- which.max(scores_sub)
  rep_index <- idx2[rep_local_pos]
  rep_name  <- colnames(U_t_mat)[rep_index]
  rep_score <- scores_sub[rep_local_pos]
  rep_global_score <- global_scores[rep_index]
  
  class_reps_excl_global <- rbind(
    class_reps_excl_global,
    data.frame(
      Class = gname,
      Index = rep_index,
      Name  = rep_name,
      WithinClassScore = rep_score,
      GlobalScore = rep_global_score,
      stringsAsFactors = FALSE
    )
  )
}

cat("Unordered class representatives:\n")
print(class_reps_excl_global)
cat("\n")
```

```{r}
# Order v2 to v6 by GLOBAL connectedness
class_reps_ordered <- class_reps_excl_global[
  order(-class_reps_excl_global$GlobalScore),
]

v2_to_v6 <- class_reps_ordered$Index

cat("Ordered v2 to v6:\n")
print(class_reps_ordered)
cat("\n")

# Full first 6 variables
first6 <- c(global_hub_index, v2_to_v6)
first6_names <- colnames(U_t_mat)[first6]

cat("v1 to v6 indices:\n")
print(first6)
cat("v1 to v6 names:\n")
print(first6_names)
cat("\n")

```
```{r}
# Order remaining assets

remaining_order <- c()

for (gname in names(groups)) {
  idx <- groups[[gname]]
  
  # remove assets already used in v1:v6
  idx_remain <- setdiff(idx, first6)
  
  if (length(idx_remain) == 0) next
  
  if (length(idx_remain) == 1) {
    remaining_order <- c(remaining_order, idx_remain)
  } else {
    tau_sub <- tau_mat[idx_remain, idx_remain, drop = FALSE]
    scores_sub <- rowSums(abs(tau_sub)) - diag(abs(tau_sub))
    
    idx_sorted <- idx_remain[order(-scores_sub)]
    remaining_order <- c(remaining_order, idx_sorted)
  }
}

remaining_names <- colnames(U_t_mat)[remaining_order]

cat("Remaining asset order:\n")
print(data.frame(
  Index = remaining_order,
  Name = remaining_names
))
cat("\n")


#  Final full C-vine ordering

c_vine_order <- c(first6, remaining_order)
c_vine_order_names <- colnames(U_t_mat)[c_vine_order]

cat("Final C-vine order (indices):\n")
print(c_vine_order)
cat("\n")

cat("Final C-vine order (names):\n")
print(c_vine_order_names)
```
```{r}




# -----------------------------------
#  Reorder PIT matrix
U_t_mat_ord <- U_t_mat[, c_vine_order]
colnames(U_t_mat_ord) <- colnames(U_t_mat)[c_vine_order]

# Optional numerical safeguard
eps <- 1e-6
U_t_mat_ord <- pmin(pmax(U_t_mat_ord, eps), 1 - eps)


# Define family set
family_set <- c(
  1,   # Gaussian
  2,   # Student t
  3,   # Clayton
  4,   # Gumbel
  5,   # Frank
  6,   # Joe
  13,  # survival Clayton
  14,  # survival Gumbel
  16   # survival Joe
)

# -----------------------------------
#  Build a C-vine structure matrix using the natural order 1:d

d <- ncol(U_t_mat_ord)

# C2RVine needs placeholder family/parameter vectors
n_pairs <- d * (d - 1) / 2

cvine_template <- C2RVine(
  order  = 1:d,
  family = rep(0, n_pairs),
  par    = rep(0, n_pairs),
  par2   = rep(0, n_pairs)
)

# Extract the structure matrix
Matrix_cvine <- cvine_template$Matrix

# -----------------------------------
# Fit the truncated C-vine
fit_cvine_vc <- RVineCopSelect(
  data          = U_t_mat_ord,
  familyset     = family_set,
  Matrix        = Matrix_cvine,
  selectioncrit = "AIC",  
  indeptest     = FALSE,
  trunclevel    = 6,
  rotations     = TRUE,
  method        = "mle",
  se            = FALSE,
  presel        = TRUE,
  cores         = 1
)

# -----------------------------------
#  Inspect the fitted model
print(fit_cvine_vc)
summary(fit_cvine_vc)
```

```{r}
saveRDS(fit_cvine_vc,"fit_cvine_vc.rds")
```

Simulations for C-vine
```{r}
set.seed(123)
```

```{r}


# USER INPUTS

W <- 500                 # rolling training window
B <- 5000                # number of simulations per step
K <- 6                   # truncation level
alpha_95 <- 0.05         # left-tail level for 95% VaR
alpha_99 <- 0.01         # left-tail level for 99% VaR
selectioncrit <- "AIC"   # or "BIC"

# VineCopula family set
# Rotations=TRUE already allows rotated/survival versions
family_set <- c(1, 2, 3, 4, 5, 6)
```

```{r}
# returns: data frame with Date column + 30 asset columns
# Reorder returns to match the C-vine variable order
asset_cols <- setdiff(names(returns), "Date")
returns_ord <- returns[, c("Date", asset_cols[c_vine_order])]

# Matrix of ordered weekly returns
ret_mat <- as.matrix(returns_ord[, -1])
dates   <- returns_ord$Date

Tn <- nrow(ret_mat)
d  <- ncol(ret_mat)


# PORTFOLIO WEIGHTS in this analysis we use the equal weights 

w <- rep(1 / d, d)

# C-VINE STRUCTURE

n_pairs <- d * (d - 1) / 2

cvine_template <- C2RVine(
  order  = 1:d,
  family = rep(0, n_pairs),
  par    = rep(0, n_pairs),
  par2   = rep(0, n_pairs)
)

Matrix_cvine <- cvine_template$Matrix


# UNIVARIATE AR(1)-GARCH(1,1)-t SPECIFICATION

spec_ar1_garch_t <- ugarchspec(
  variance.model = list(model = "sGARCH", garchOrder = c(1, 1)),
  mean.model     = list(armaOrder = c(1, 0), include.mean = TRUE),
  distribution.model = "std"
)


# CONTAINERS FOR ROLLING BACKTEST

n_oos <- Tn - W

VaR_95 <- rep(NA_real_, n_oos)
VaR_99 <- rep(NA_real_, n_oos)

realized_port <- rep(NA_real_, n_oos)
viol_95       <- rep(NA_integer_, n_oos)
viol_99       <- rep(NA_integer_, n_oos)

forecast_dates <- dates[(W + 1):Tn]
fit_success    <- rep(TRUE, n_oos)

# storage for diagnostics
cvine_fits <- vector("list", n_oos)


#  ROLLING BACKTEST LOOP

for (s in 1:n_oos) {
  
  cat("Rolling step", s, "of", n_oos, "\n")
  
  
  # Rolling train/test split
  train_idx <- s:(s + W - 1)
  test_idx  <- s + W
  
  R_train <- ret_mat[train_idx, , drop = FALSE]
  R_test  <- ret_mat[test_idx, ]
  
  # Containers for this rolling window
  U_train  <- matrix(NA_real_, nrow = W, ncol = d)
  mu_fc    <- rep(NA_real_, d)
  sigma_fc <- rep(NA_real_, d)
  shape_fc <- rep(NA_real_, d)
  
  colnames(U_train) <- colnames(R_train) <- colnames(ret_mat)
  
  step_ok <- TRUE
  
  
  # Fit AR(1)-GARCH(1,1)-t marginals and compute PITs
  for (j in 1:d) {
    
    x <- R_train[, j]
    
    fit_j <- tryCatch(
      ugarchfit(
        spec   = spec_ar1_garch_t,
        data   = x,
        solver = "hybrid"
      ),
      error = function(e) {
        tryCatch(
          ugarchfit(
            spec   = spec_ar1_garch_t,
            data   = x,
            solver = "solnp"
          ),
          error = function(e2) NULL
        )
      }
    )
    
    if (is.null(fit_j)) {
      cat("GARCH failed for", colnames(ret_mat)[j], "at step", s, "\n")
      step_ok <- FALSE
      break
    }
    
    # Standardized residuals
    z_j <- residuals(fit_j, standardize = TRUE)
    
    # Student-t shape parameter
    coef_j <- coef(fit_j)
    shape_j <- unname(coef_j["shape"])
    
    if (is.na(shape_j) || shape_j <= 2) {
      cat("Invalid shape parameter for", colnames(ret_mat)[j], "at step", s, "\n")
      step_ok <- FALSE
      break
    }
    
    shape_fc[j] <- shape_j
    
    # Parametric PIT
    U_train[, j] <- pdist(
      distribution = "std",
      q            = as.numeric(z_j),
      mu           = 0,
      sigma        = 1,
      shape        = shape_j
    )
    
    # Clip PITs slightly away from 0 and 1
    eps <- 1e-6
    U_train[, j] <- pmin(pmax(U_train[, j], eps), 1 - eps)
    
    # One-step-ahead forecast
    fc_j <- tryCatch(
      ugarchforecast(fit_j, n.ahead = 1),
      error = function(e) NULL
    )
    
    if (is.null(fc_j)) {
      cat("Forecast failed for", colnames(ret_mat)[j], "at step", s, "\n")
      step_ok <- FALSE
      break
    }
    
    mu_fc[j]    <- as.numeric(fitted(fc_j))
    sigma_fc[j] <- as.numeric(sigma(fc_j))
    
    if (is.na(mu_fc[j]) || is.na(sigma_fc[j]) || sigma_fc[j] <= 0) {
      cat("Invalid forecast for", colnames(ret_mat)[j], "at step", s, "\n")
      step_ok <- FALSE
      break
    }
  }
  
  if (!step_ok) {
    fit_success[s] <- FALSE
    next
  }
  
  
  # Fit truncated C-vine copula on PITs
  fit_cvine_s <- tryCatch(
    RVineCopSelect(
      data          = U_train,
      familyset     = family_set,
      Matrix        = Matrix_cvine,
      selectioncrit = selectioncrit,
      indeptest     = FALSE,
      trunclevel    = K,
      rotations     = TRUE,
      method        = "mle",
      se            = FALSE,
      presel        = TRUE,
      cores         = 1
    ),
    error = function(e) NULL
  )
  
  if (is.null(fit_cvine_s)) {
    cat("C-vine fit failed at step", s, "\n")
    fit_success[s] <- FALSE
    next
  }
  
  cvine_fits[[s]] <- fit_cvine_s
  
  
  # Simulate dependent uniforms from fitted C-vine
  U_sim <- tryCatch(
    RVineSim(B, fit_cvine_s),
    error = function(e) NULL
  )
  
  if (is.null(U_sim)) {
    cat("C-vine simulation failed at step", s, "\n")
    fit_success[s] <- FALSE
    next
  }
  
  
  # Map uniforms -> t innovations -> return simulations
  R_sim <- matrix(NA_real_, nrow = B, ncol = d)
  
  for (j in 1:d) {
    z_sim_j <- qdist(
      distribution = "std",
      p            = U_sim[, j],
      mu           = 0,
      sigma        = 1,
      shape        = shape_fc[j]
    )
    
    R_sim[, j] <- mu_fc[j] + sigma_fc[j] * z_sim_j
  }
  
 
  # Simulated portfolio returns and VaR forecasts
  port_sim <- as.numeric(R_sim %*% w)
  
  VaR_95[s] <- as.numeric(quantile(port_sim, probs = alpha_95, na.rm = TRUE))
  VaR_99[s] <- as.numeric(quantile(port_sim, probs = alpha_99, na.rm = TRUE))
  
  # Realized next-week portfolio return
  realized_port[s] <- sum(R_test * w)
  
  # VaR violations
  viol_95[s] <- as.integer(realized_port[s] < VaR_95[s])
  viol_99[s] <- as.integer(realized_port[s] < VaR_99[s])
}


# BACKTEST RESULTS DATA FRAM

backtest_cvine <- data.frame(
  Date       = forecast_dates,
  Realized   = realized_port,
  VaR_95     = VaR_95,
  VaR_99     = VaR_99,
  Viol_95    = viol_95,
  Viol_99    = viol_99,
  FitSuccess = fit_success
)

print(head(backtest_cvine))
print(tail(backtest_cvine))
```
```{r}
saveRDS(backtest_cvine,"backtest_cvine.rds")
```

```{r}
readRDS("backtest_cvine.rds")
```

```{r}

# SUMMARY STATISTICS


violation_rate_95 <- mean(backtest_cvine$Viol_95[backtest_cvine$FitSuccess], na.rm = TRUE)
violation_rate_99 <- mean(backtest_cvine$Viol_99[backtest_cvine$FitSuccess], na.rm = TRUE)

cat("95% VaR violation rate =", violation_rate_95, "\n")
cat("99% VaR violation rate =", violation_rate_99, "\n")

cat("95% number of violations =", sum(backtest_cvine$Viol_95, na.rm = TRUE), "\n")
cat("99% number of violations =", sum(backtest_cvine$Viol_99, na.rm = TRUE), "\n")

cat("Number of successful fits =", sum(backtest_cvine$FitSuccess), "\n")
cat("Number of failed fits =", sum(!backtest_cvine$FitSuccess), "\n")
```

```{r}

#  KUPIEC UNCONDITIONAL COVERAGE TEST

kupiec_test <- function(violations, alpha) {
  x <- sum(violations, na.rm = TRUE)
  n <- sum(!is.na(violations))
  
  if (n == 0) {
    return(data.frame(
      alpha = alpha,
      n = 0,
      x = NA,
      hit_rate = NA,
      LRuc = NA,
      p_value = NA
    ))
  }
  
  p_hat <- x / n
  
  if (p_hat == 0 || p_hat == 1) {
    return(data.frame(
      alpha = alpha,
      n = n,
      x = x,
      hit_rate = p_hat,
      LRuc = NA,
      p_value = NA
    ))
  }
  
  LRuc <- -2 * (
    (n - x) * log((1 - alpha) / (1 - p_hat)) +
      x     * log(alpha / p_hat)
  )
  
  p_value <- 1 - pchisq(LRuc, df = 1)
  
  data.frame(
    alpha = alpha,
    n = n,
    x = x,
    hit_rate = p_hat,
    LRuc = LRuc,
    p_value = p_value
  )
}

kupiec_95 <- kupiec_test(
  backtest_cvine$Viol_95[backtest_cvine$FitSuccess],
  alpha = alpha_95
)

kupiec_99 <- kupiec_test(
  backtest_cvine$Viol_99[backtest_cvine$FitSuccess],
  alpha = alpha_99
)

kupiec_results_cvine <- rbind(kupiec_95, kupiec_99)
print(kupiec_results_cvine)
```





```{r}
#PLOTTING THE RESULTS 
bt_cv <- subset(backtest_cvine, FitSuccess == TRUE)

plot(
  bt_cv$Date, bt_cv$Realized, type = "l",
  main = "Realized Portfolio Returns vs VaR Forecasts:Truncated C-Vine",
  xlab = "Date", ylab = "Realized Returns"
)

lines(bt_cv$Date, -abs(bt_cv$VaR_95), lty = 1, lwd = 1.5,col = "blue")
lines(bt_cv$Date, -abs(bt_cv$VaR_99), lty = 1, lwd = 1.5,col = "red")

legend(
  "bottomleft",
  legend = c("Realized portfolio return", "-VaR 95%", "-VaR 99%"),
  lty = c(1, 1, 1),
  col = c("black", "blue", "red"),
  bty = "n"
)
```





We also add the rolling MV portfolio for backtesting  as a baseline model 
```{r}

# Prepare data
returns <- returns %>% arrange(Date)
asset_cols <- setdiff(names(returns), "Date")


#  Helper function:long-only global minimum-variance weights

minvar_weights_longonly <- function(Sigma) {
  d <- ncol(Sigma)
  
  # quadratic term
  Dmat <- 2 * as.matrix(Sigma)
  Dmat <- (Dmat + t(Dmat)) / 2
  
  # add tiny ridge for numerical stability
  Dmat <- Dmat + diag(1e-8, d)
  
  dvec <- rep(0, d)
  
  # Constraints for solve.QP:
  # Amat' w >= bvec
  #
  # First column: sum(w) = 1
  # Remaining columns: w_i >= 0
  Amat <- cbind(rep(1, d), diag(d))
  bvec <- c(1, rep(0, d))
  
  sol <- solve.QP(
    Dmat = Dmat,
    dvec = dvec,
    Amat = Amat,
    bvec = bvec,
    meq = 1
  )
  
  w <- sol$solution
  names(w) <- colnames(Sigma)
  return(w)
}


# Rolling-window settings

W <- 500
Tn <- nrow(returns)

# store results
backtest_mv_norm <- data.frame(
  Date = returns$Date[(W + 1):Tn],
  VaR_95 = NA_real_,
  VaR_99 = NA_real_,
  Rp_real = NA_real_,
  mu_p = NA_real_,
  sigma_p = NA_real_,
  Violation_95 = NA_integer_,
  Violation_99 = NA_integer_
)

# store weights from each rolling step
weights_list <- vector("list", Tn - W)


# 3. Rolling backtest loop
for (tt in W:(Tn - 1)) {
  
  cat("Rolling MV normal VaR step:", tt, "of", Tn - 1, "\n")
  
  #----------------------------------------------
  # Define training and test samples
  #----------------------------------------------
  train_idx <- (tt - W + 1):tt
  test_idx  <- tt + 1
  
  train_data <- returns[train_idx, ]
  test_data  <- returns[test_idx, ]
  
  # matrix of training returns (W x d)
  R_train <- as.matrix(train_data[, asset_cols])
  
  #----------------------------------------------
  # Estimate sample mean vector and covariance
  #----------------------------------------------
  mu_hat <- colMeans(R_train, na.rm = TRUE)
  Sigma_hat <- cov(R_train, use = "pairwise.complete.obs")
  
  #----------------------------------------------
  # Compute rolling long-only minimum-variance weights
  #----------------------------------------------
  w_t <- tryCatch(
    minvar_weights_longonly(Sigma_hat),
    error = function(e) NULL
  )
  
  if (is.null(w_t)) {
    cat("MV optimization failed at step", tt, "\n")
    next
  }
  
  weights_list[[tt - W + 1]] <- w_t
  
  #----------------------------------------------
  # Portfolio mean and standard deviation
  #----------------------------------------------
  mu_p_t <- as.numeric(sum(w_t * mu_hat))
  sigma_p_t <- as.numeric(sqrt(t(w_t) %*% Sigma_hat %*% w_t))
  
  #----------------------------------------------
  # Delta-normal / variance-covariance VaR
  # qnorm(0.05) < 0 and qnorm(0.01) < 0
  # VaR reported as positive loss number
  #----------------------------------------------
  VaR_95_t <- -(mu_p_t + qnorm(0.05) * sigma_p_t)
  VaR_99_t <- -(mu_p_t + qnorm(0.01) * sigma_p_t)
  
  #----------------------------------------------
  # F. Realized next-period portfolio return: using the same rolling weights
  #----------------------------------------------
  R_test <- as.numeric(test_data[1, asset_cols])
  Rp_real_t <- sum(w_t * R_test)
  
  #----------------------------------------------
  # VaR violation indicators
  #----------------------------------------------
  viol_95_t <- as.integer(Rp_real_t < -VaR_95_t)
  viol_99_t <- as.integer(Rp_real_t < -VaR_99_t)
  
  #----------------------------------------------
  # Save results
  #----------------------------------------------
  row_out <- tt - W + 1
  
  backtest_mv_norm$VaR_95[row_out] <- VaR_95_t
  backtest_mv_norm$VaR_99[row_out] <- VaR_99_t
  backtest_mv_norm$Rp_real[row_out] <- Rp_real_t
  backtest_mv_norm$mu_p[row_out] <- mu_p_t
  backtest_mv_norm$sigma_p[row_out] <- sigma_p_t
  backtest_mv_norm$Violation_95[row_out] <- viol_95_t
  backtest_mv_norm$Violation_99[row_out] <- viol_99_t
}


```

```{r}
bt_mv_norm <- backtest_mv_norm %>%
  filter(!is.na(VaR_95), !is.na(Rp_real))

nrow(bt_mv_norm)

mean(bt_mv_norm$Violation_95)
mean(bt_mv_norm$Violation_99)

sum(bt_mv_norm$Violation_95)
sum(bt_mv_norm$Violation_99)

```
```{r}
kupiec_test <- function(violations, alpha) {
  x <- sum(violations, na.rm = TRUE)
  n <- sum(!is.na(violations))
  p_hat <- x / n
  
  if (p_hat == 0 || p_hat == 1) {
    return(list(
      LRuc = NA,
      p_value = NA,
      violations = x,
      n = n,
      observed_rate = p_hat
    ))
  }
  
  LRuc <- -2 * (
    (n - x) * log(1 - alpha) + x * log(alpha) -
      ((n - x) * log(1 - p_hat) + x * log(p_hat))
  )
  
  p_value <- 1 - pchisq(LRuc, df = 1)
  
  list(
    LRuc = LRuc,
    p_value = p_value,
    violations = x,
    n = n,
    observed_rate = p_hat
  )
}

kupiec_95 <- kupiec_test(bt_mv_norm$Violation_95, alpha = 0.05)
kupiec_99 <- kupiec_test(bt_mv_norm$Violation_99, alpha = 0.01)

kupiec_95
kupiec_99

```

```{r}
#PLOT THE MV GRAPH
plot(
  bt_mv_norm$Date, bt_mv_norm$Rp_real, type = "l",
  main = "Realized Portfolio Returns vs VaR Forecasts:Rolling MV Portfolio",
  xlab = "Date", ylab = "Realized Returns"
)

lines(bt_mv_norm$Date, -bt_mv_norm$VaR_95, lty = 1, lwd = 1.5,col = "blue")
lines(bt_mv_norm$Date, -bt_mv_norm$VaR_99, lty = 1, lwd = 1.5,col = "red")

legend(
  "bottomleft",
  legend = c("Realized MV return", "-VaR 95%", "-VaR 99%"),
  lty = c(1, 1, 1),
  col = c("black", "blue", "red"),
  bty = "n"
)

```
